### 2020-07-22 eda.Rmd
Add .Rproj and recreate Table II

---
output:
  pdf_document
fontsize: 10pt 
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(dev = "cairo_pdf")
knitr::opts_chunk$set(fig.width = 6)
knitr::opts_chunk$set(fig.height = 4)

options(knitr.kable.NA = "")
options(scipen = 999)

library(tidyverse)
library(haven)

```

```{r}
seed011204 <-read_stata("dataverse_files/seedanalysis_011204_1.dta")

seed080404 <- read_stata("dataverse_files/seedanalysis_080404_1.dta")

combined <- read_stata("dataverse_files/seedanalysis_011204_080404_1.dta")

```

```{r}

combined <- combined  %>%
  # turn dollars into hundreds of dollars
  mutate(
    totbal = totbal / 100,
    newtotbal = newtotbal / 100
  ) %>%
  # fix population variable
  mutate(pop = as.numeric(str_replace(pop, ",", ""))) %>%
  # create bank penetration variable%>%
  mutate(brgy_penetration = no_clients / pop)

combined <- combined %>%
  group_by(brgy_penetration) %>%
  mutate(sd_totbal = sd(totbal),
         mean_totbal = mean(totbal)) %>%
  ungroup()

treated <- combined %>%
  filter(!is.na(treatment))

treated <- treated %>%
  # create control dummy
  mutate(control = as.numeric(group == "C")) %>%
  # create distance to bank variable
  mutate(dist_GB = 
           case_when(
             butuan == 1 ~ dbutuan,
             ampayon == 1 ~ dampayon
           )
  )

treated %>%
  group_by(group) %>%
  summarize(n(), mean(totbal), sd(totbal))

```

## Remake Table II

### Panel A.

```{r}
summary(lm(totbal ~ factor(group) - 1, data = treated))
summary(aov(totbal ~ factor(group), data = treated))

summary(lm(active ~ factor(group) - 1, data = treated))
summary(aov(active ~ factor(group), data = treated))

summary(lm(dist_GB ~ factor(group) - 1, data = treated))
summary(aov(dist_GB ~ factor(group), data = treated))

summary(lm(brgy_penetration ~ factor(group) - 1, data = treated))
summary(aov(brgy_penetration ~ factor(group), data = treated))

summary(lm(sd_totbal ~ factor(group) - 1, data = treated))
summary(aov(sd_totbal ~ factor(group), data = treated))

summary(lm(mean_totbal ~ factor(group) - 1, data = treated))
summary(aov(mean_totbal ~ factor(group), data = treated))

summary(lm(pop ~ factor(group) - 1, data = treated))
summary(aov(mean_totbal ~ factor(group), data = treated))
```

### Panel B.

```{r}
summary(lm(yearsed ~ factor(group) - 1, data = treated))
summary(aov(yearsed ~ factor(group), data = treated))

summary(lm(female ~ factor(group) - 1, data = treated))
summary(aov(female ~ factor(group), data = treated))

summary(lm(age ~ factor(group) - 1, data = treated))
summary(aov(age ~ factor(group), data = treated))

summary(lm(impatient_mon01 ~ factor(group) - 1, data = treated))
summary(aov(impatient_mon01 ~ factor(group), data = treated))

summary(lm(hyper_mon_new2 ~ factor(group) - 1, data = treated))
summary(aov(hyper_mon_new2 ~ factor(group), data = treated))

count(treated, group)
```

Questions:

1. We examine whether individuals who exhibit hyperbolic preferences in hypothetical time preference questions are more likely to open such accounts, since theoretically these individuals may have a preference for commitment. 
2. Second, we test whether such individuals save more as a result of opening the account.

Our questions:

* Remake table 2 with non-parametric standard errors
* Remake table 3 (meh)
* Remake table 6
* Remake table 7

* Did the respondents make positive savings?
* What proportion of people opened the account?
* Did the treatment group save more than the control group?
* estimate the pdf of savings and then estimate the percentiles

```{r}
count(combined, group)

```