### 2020-07-22 eda.Rmd
Remake Table VII

---
output:
  pdf_document
fontsize: 10pt 
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(dev = "cairo_pdf")
knitr::opts_chunk$set(fig.width = 6)
knitr::opts_chunk$set(fig.height = 4)

options(knitr.kable.NA = "")
options(scipen = 999)

library(tidyverse)
library(haven)
library(quantreg)
```

```{r}
seed011204 <-read_stata("dataverse_files/seedanalysis_011204_1.dta")

seed080404 <- read_stata("dataverse_files/seedanalysis_080404_1.dta")

combined <- read_stata("dataverse_files/seedanalysis_011204_080404_1.dta")

```

```{r}
combined <- combined  %>%
  # turn dollars into hundreds of dollars
  mutate(
    totbal = totbal / 100,
    newtotbal = newtotbal / 100
  ) %>%
  # fix population variable
  mutate(pop = as.numeric(str_replace(pop, ",", ""))) %>%
  # create bank penetration variable%>%
  mutate(brgy_penetration = no_clients / pop)

combined <- combined %>%
  # calculate the mean and sd for each bank
  # the Stata code is a little cyptic
  group_by(brgy_penetration) %>%
  mutate(sd_totbal = sd(totbal),
         mean_totbal = mean(totbal)) %>%
  ungroup()

# create a set with the 1777 observations that are in the SEED, 
# marketing treatment, or control group
treated <- combined %>%
  filter(!is.na(treatment))

treated <- treated %>%
  # create control dummy
  mutate(control = as.numeric(group == "C")) %>%
  # create distance to bank variable
  mutate(dist_GB = 
           case_when(
             butuan == 1 ~ dbutuan,
             ampayon == 1 ~ dampayon
           )
  )

treated %>%
  group_by(group) %>%
  summarize(n(), mean(totbal), sd(totbal))

```

## Remake Table II

### Panel A.

```{r}
summary(lm(totbal ~ factor(group) - 1, data = treated))
summary(aov(totbal ~ factor(group), data = treated))

summary(lm(active ~ factor(group) - 1, data = treated))
summary(aov(active ~ factor(group), data = treated))

summary(lm(dist_GB ~ factor(group) - 1, data = treated))
summary(aov(dist_GB ~ factor(group), data = treated))

summary(lm(brgy_penetration ~ factor(group) - 1, data = treated))
summary(aov(brgy_penetration ~ factor(group), data = treated))

summary(lm(sd_totbal ~ factor(group) - 1, data = treated))
summary(aov(sd_totbal ~ factor(group), data = treated))

summary(lm(mean_totbal ~ factor(group) - 1, data = treated))
summary(aov(mean_totbal ~ factor(group), data = treated))

summary(lm(pop ~ factor(group) - 1, data = treated))
summary(aov(mean_totbal ~ factor(group), data = treated))
```

### Panel B.

```{r}
summary(lm(yearsed ~ factor(group) - 1, data = treated))
summary(aov(yearsed ~ factor(group), data = treated))

summary(lm(female ~ factor(group) - 1, data = treated))
summary(aov(female ~ factor(group), data = treated))

summary(lm(age ~ factor(group) - 1, data = treated))
summary(aov(age ~ factor(group), data = treated))

summary(lm(impatient_mon01 ~ factor(group) - 1, data = treated))
summary(aov(impatient_mon01 ~ factor(group), data = treated))

summary(lm(hyper_mon_new2 ~ factor(group) - 1, data = treated))
summary(aov(hyper_mon_new2 ~ factor(group), data = treated))

count(treated, group)
```

## Remake Table VII

```{r}
rq(as.numeric(balchange) ~ treatment + marketing, 
   tau = seq(from = 0.1, to = 0.9, by = 0.1),
   data = filter(seed011204, !is.na(treatment)))

rq(as.numeric(balchange) ~ treatment + marketing, 
   tau = seq(from = 0.1, to = 0.9, by = 0.1),
   data = treated)
```

## Questions

* Remake table 6 with bootstrapped standard errors
* Remake table 7 with kernel density estimators and compare with the percentiles percentiles from the quantile regression
* Estimate the proportion of people who would open an account
* Do a k-sample test to compare the savings or change in svaings of the three groups. Then do pair-wise tests with corrections for multiple testing. 
* We can use LOESS/KDE to compare any of the bivariate relationships
    * Year of education by totbal or change in totbal
    * Age or age2 by totbal or change in totbal

1. We examine whether individuals who exhibit hyperbolic preferences in hypothetical time preference questions are more likely to open such accounts, since theoretically these individuals may have a preference for commitment. 
2. Second, we test whether such individuals save more as a result of opening the account.